---
title: "Introduction to R - Solutions"
author: "Simon Munzert"
date: "18.09.2017"
output: html_document
---


```{r setup, include=FALSE}
source("packages.r")
```

<br>

#### In-class exercise 

##### 1. How many flights started from all the NYC airports in 2013?

```{r}
nrow(flights)
```

##### 2. How many flights started from NYC airports in 2013, by month?

```{r}
table(flights$month)
flights_by_month <- group_by(flights, month)
summarize(flights_by_month, count = n())
```

##### 3. Which were the most frequent destinations?

```{r}
flights_by_dest <- group_by(flights, dest) 
dat <- summarize(flights_by_dest, n_per_dest = n())
arrange(dat, desc(n_per_dest))
# alternative: 
# sort(table(flights$dest))
```

##### 4. Which flights went from JFK to PWM (Portland) in January between midnight and 5am?

```{r}
dat <- filter(flights, origin == "JFK", month == 1, dest == "PWM", dep_time < 500) 
dat
```

##### 5. Which carrier had the biggest mean net delay (departure delay minus arrival delay), which the lowest?

```{r}
dat <- group_by(flights, carrier)
dat <- mutate(dat, net_delay = dep_delay - arr_delay)
dat_sum <- summarize(dat, delay_mean = mean(net_delay, na.rm= TRUE))
arrange(dat_sum, delay_mean) 
```

##### 6. Inspect visually how the net delays are distributed!

```{r}
hist(dat$net_delay)
```

##### 7. How are distance and net delay related?

```{r}
summary(lm(net_delay ~ distance, data = dat))
with(dat, cor(net_delay, distance))
```


<br>

#### Working with vectors 


##### 1. Create a vector x with elements [0,4,8,12,16,20].

```{r}
x <- seq(0, 20, 4)
x
```

##### 2. Create a vector y with elements [3,3,3,4,4,4,4,5,5,5,5,5].

```{r}
y <- c(3,3,3,4,4,4,4,5,5,5,5,5)
rep(c(3, 4, 5), c(3, 4, 5))
```

##### 3. Combine the first five elements of x with elements 2 to 12 of vector y to create a new vector z.

```{r}
z <- c(x[1:5], y[2:12])
```

##### 4. What's the sum of all numbers between 1 and 100? Use the functions you learned in the vectors script to compute it!

```{r}
sum(seq(1, 100))
sum(1:100)
```

##### 5. What's the sum of all odd numbers between 1 and 100 squared?

```{r}
sum(seq(1, 100, 2)^2)
```


<br>

#### Working with data frames 

##### 1. Inspect the "babynames" package! What data does it contain? Focus on the babynames data frame in the following.

```{r}
names(babynames)
nrow(babynames)
```

##### 2. Which time period does the data frame cover?

```{r}
unique(babynames$year)
range(babynames$year)
```

##### 3. What has been the most frequent name in the US since 1880? (hint: you might need the group_by() as well as the sum() function)

```{r}
babynames_byname <- group_by(babynames, name)
babynames_byname_sum <- summarize(babynames_byname, total = sum(n))
head(arrange(babynames_byname_sum, desc(total)))
```

##### 4. Use graphical means to determine since which year the name "Kim" became more popular as a female than a male name! (hint: grouping by two variables is helpful here! For the plot, use "year" on the x axis and "total count" on the y axis! Furthermore, this Stackoverflow post helps you understand how to plot two lines in one chart: https://stackoverflow.com/questions/2564258/plot-two-graphs-in-same-plot-in-r)

```{r}
kim_df <- filter(babynames, name == "Kim")
#kim_df <- babynames[babynames$name == "Kim",]
kim_grouped <- group_by(kim_df, year, sex)
kim_sum <- summarize(kim_grouped, total = sum(n))
plot(kim_sum$year[kim_sum$sex == "F"], kim_sum$total[kim_sum$sex == "F"], type = "l")
lines(kim_sum$year[kim_sum$sex == "M"], kim_sum$total[kim_sum$sex == "M"], col = "red")
qplot(year, total, color = sex, data = kim_sum, geom = "line") # ggplot-style alternative
```









