---
title: "Interactions - Assignment"
author: "Simon Munzert"
date: "21.09.2017"
output: html_document
---


```{r setup, include=FALSE}
source("packages.r")
```

<br>

### Assignment

#### 1. Replicate the results from the motivating setup Braumoeller (2004, IO 58) uses to make his point and that leads to Table 1. Use the parameters from footnote 1! 

```{r}
# model 1
set.seed(1)
N <- 1000
x1 <- runif(N, 0, 1)
x2 <- runif(N, 0, 1)
e <- rnorm(N, 0, 1)
y <- 0.2 + x1 + 2*x2 + 3*x1*x2 + e
summary(model1 <- lm(y ~ x1 + x2 + x1*x2))
```

```{r}
# model 2: recoding x2
x2_star <- x2 + 0.4
y <- 0.2 + x1 + 2*x2_star + 3*x1*x2_star + e
summary(model2 <- lm(y ~ x1 + x2_star + x1*x2_star))
stargazer(model1, model2, header = FALSE, style = "apsr", type = "html", out = "../output/braumoeller_replication.html")
```


<br>

#### 2. Consider the "wage2" data set from the wooldridge package! The outcome of interest is monthly earnings. Make yourself familiar with the data set and, based on the observational data, come up with ad hoc hypotheses that predict wage. Also, try to develop a hypothesis where one variable (continuous) mediates the effect of another. Estimate a model that predicts wage using a set of variables, including the interaction variables. Visualize the interaction effect appropriately and interpret your findings!

```{r}
data(wage2)
summary(model_out <- lm(wage ~ tenure + educ + exper*IQ, data = wage2)) # IQ probably more important when experience is low?

interplot(m = model_out, var1 = "exper", var2 = "IQ", hist  = TRUE) +
  xlab('Years of work experience') +
  ylab('Estimated coefficient for IQ') +
  ggtitle('Estimated coefficient of IQ\non wage by years of experience') +
  theme(plot.title = element_text(face='bold'))
```


<br>

#### 3. What determines support for European integration? For the following tasks, please consider the data set "eu-support.dta", which is a small dataset on individual support for European integration (variable eusupp) in Denmark, Spain and the UK. Your preliminary theorizing suggests that EU support is determined by political ideology and by the country-specific understanding of EU integration.

##### a) Load your data. Next, recode the variable "nation" as follows: 7 = Denmark, 9 = UK, 11 = Spain using the recode() function from the car package. Finally, make the variable a character variable using as.character().

```{r}
dat <- read_dta("../data/eu-support.dta")
table(dat$nation)
dat$nation <- car::recode(dat$nation, "7 = 'Denmark' ; 9 = 'UK' ; 11 = 'Spain'") 
dat$nation <- as.factor(dat$nation)
table(dat$nation)
```

##### b) Build a simple model that predicts EU support with left-right ideology and the country of residence as well as their interactions. Also, control for gender (2 is female), age, and education. Display the results in a regression table!

```{r}
summary(m1 <- lm(eusupp ~ nation*left_right + sex + age + educ, data = dat))
```

##### c) Try to interpret the effect of ideology across the three countries. What are the partial slopes (the effect of ideology on EU support) in each country?

```{r}
b_dk <- coef(m1)['left_right']
b_es <- coef(m1)['left_right'] + coef(m1)['nationSpain:left_right']
b_uk <- coef(m1)['left_right'] + coef(m1)['nationUK:left_right']
b_dk
b_es
b_uk
```

##### d) Test whether these slopes are statistically significantly different from 0. Hint: applying the vcov() command on your model gives you the variance-covariance matrix!

```{r}
# retrieve variances and covariances of the coefficients
var_b_lr <- vcov(m1)[4,4]
var_b_lrXuk <- vcov(m1)[7,7]
var_b_lrXes <- vcov(m1)[8,8]
cov_b_lr_lrXuk <- vcov(m1)[4,7]
cov_b_lr_lrXes <- vcov(m1)[4,8]
```

```{r}
# calculate the standard errors of the conditional effects  
se_b_es <- sqrt(var_b_lr + var_b_lrXes + 2 * cov_b_lr_lrXes)
se_b_uk <- sqrt(var_b_lr + var_b_lrXuk + 2 * cov_b_lr_lrXuk)  
```

```{r}
# obtain the p-value for the conditional effect in Spain
pnorm(q = abs((b_es / se_b_es)), mean = 0, sd = 1, lower.tail = F)*2
```

```{r}
# obtain the p-value for the conditional effect in the UK
pnorm(q = abs((b_uk / se_b_uk)), mean = 0 , sd = 1, lower.tail = F)*2
```


##### e) Compute 95% confidence intervals for the effects and visualize the effects (along with the intervals) in a plot!

```{r}
# compute confidence intervals
conf95_denmark <- confint(m1, "left_right", level = .95)
conf95_spain <- b_es + 2* se_b_es * c(-1, 1)
conf95_uk <- b_uk + 2* se_b_uk * c(-1, 1)
```

```{r}
# plot
#pdf(file="../output/effectsplot.pdf", height=6, width=4, family="URWTimes")
par(las = 1) # horizontal axis labels
nations <- c("Denmark", "Spain", "UK")
effects <- c(b_dk, b_es, b_uk)
# plot point estimates
plot(seq_along(nations), effects,  pch = 19, xlim = c(.5, 3.5), ylim = c(-.15, .2), xaxt = "n", xaxt = "n", ylab = "Estimated effect of left-right position", xlab = "")
# add 0 line
abline(h = 0, lty = 2)
# add nations labels
axis(1, seq_along(nations), nations, tick = FALSE)
# add confidence intervals
segments(1, conf95_denmark[1], 1, conf95_denmark[2])
segments(2, conf95_spain[1], 2, conf95_spain[2])
segments(3, conf95_uk[1], 3, conf95_uk[2])
#dev.off()
```

#### 4. (Optional): check out how the "margins" package works! A good start might be
browseURL("https://cran.r-project.org/web/packages/margins/vignettes/Introduction.html")
# Try to understand how this package is not only useful for visualizing conditional effects, but to calculate marginal effects in general and how this helps you communicate the implications of your models better!
